The impact of cosmic variance on simulating weak lensing surveys

### 2014-10-07 Writeup/paper_v3.tex
significant editing in 4.1.2

To quantify galaxy morphology and shape, we will use several
quantities from the above fits.  First, from the single \sersic
profile fits, we use the \sersic index and the axis ratio.  The axis
ratio can also be used to derive a distortion,
\begin{equation}\label{eq:distortion}
e = \frac{1-q^2}{1+q^2}
\end{equation}
or an ellipticity,
\begin{equation}
\varepsilon = \frac{1-q}{1+q}.
\end{equation}

@@

We can test the influence of environment on the galaxy shapes by
comparing the distributions of the axis ratios for the overdense and
underdense redshift slices, or by encapsulating that distribution as a
single number, the RMS ellipticity. By volume-limiting the sample, we can avoid issues
wherein the flux limit leads to artificial changes in the sample as a
function of redshift.  We will also carry out tests to differentiate
between environmental effects versus evolution of the population with
redshift (at fixed mass).

%Then, we proceed to consider the root mean square ellipticity, a statistic that can characterize the shape of the population/sample of galaxies. 
\subsubsection{Comparing distributions}
\begin{figure*}
 \centering
 \includegraphics[width=2.2\columnwidth]{axis_ratio_all}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_Bbandevolution.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_masscut.eps}
 \caption{The distributions of axis ratios of galaxies in \emph{all}
   overdense (OD) and \emph{all} underdense (UD) regions in the case
   of the luminosity-selected sample \s1 (left), luminosity-selected
   sample with $B$-band evolution taken into account \s2 (center), and
   thestellar-mass-selected sample \s3 (right). The upper panels show
   the histograms, 
 and the bottom panels show the cumulative distribution functions
 (CDF). The $p$-values computed using these CDFs are shown in Table. \ref{table:pvalues_all}.}
 \label{fig:axisratio_all}
\end{figure*}

@@

\subsubsection{RMS ellipticites}
\begin{figure}
 %\centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip1_noevolution.eps} \
 \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \caption{Left: RMS ellipticities with ellipticity defined as $\frac{1-q}{1+q}$. \; 
          Right: RMS ellipticities with ellipticity defined as $\frac{1-q^2}{1+q^2}$.\\ The horizontal errorbars simply correspond to the binwidth while the vertical ones
          are $1\sigma$ errorbars obtained by bootstrapping. \rachel{I'm thinking these are so similar we just need to show one figure, and can say the results for the other in words.  The same goes for the next figure.}}
 \label{fig:rms_ellip}
\end{figure}

\begin{figure}
 %\centering
  \includegraphics[width=1.0\columnwidth]{rms_ellip1_Bbandevolution.eps} \
\includegraphics[width=1.0\columnwidth]{rms_ellip2_Bbandevolution.eps} \\
 \caption{Left: RMS ellipticities with ellipticity defined as $\frac{1-q}{1+q}$. \; 
          Right: RMS ellipticities with ellipticity defined as $\frac{1-q^2}{1+q^2}$.\\ The horizontal errorbars simply correspond to the binwidth while the vertical ones
          are $1\sigma$ errorbars obtained by bootstrapping. The solid points correspond to the sample where the luminosity cut evolves by -1.23 per unit redshift (\s$2$) and the unfilled
          points correspond to the sample obtained from stellar mass cuts (\s$3$).}
 \label{fig:rms_ellip_Bband_mass}
\end{figure}

\begin{figure}
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_noevolution.eps} \
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_Bbandevolution_masscut.eps} \\
 \caption{RMS ellipticies with ellipticity defined as $\sqrt{e_1^2+e_2^2}$, where $e_1$ and $e_2$ are the real and imaginary components of the distortion defined in Eq.~\ref{eq:distortion}.
          The horizontal errorbars simply correspond to the binwidth while the vertical ones
          are $1\sigma$ errorbars obtained by bootstrapping. The solid points correspond to the luminosity selected sample (\s$1$ sample on the left and \s$2$ sample on the right) and the unfilled
          points correspond to the sample obtained from stellar mass cuts (\s$3$). }
 \label{fig:rms_ellip_momentbased}         
\end{figure}

%The shape of a population of galaxies can be characterized by a single number called `RMS ellipticity'. If $q=b/a$ is the axis-ratio, then the ellipticity maybe defined as $\frac{1-q}{1+q}$
%or as $\frac{1-q^2}{1+q^2}$. The root mean squared (RMS) of ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}.
For the luminosity-selected sample without any evolution, RMS ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}. As one can see, the underdense regions
have significantly higher values for RMS ellipticities when compared to the overdense regions. Note in particular that we've been able to capture the narrow overdense bin $0.75\le z < 0.80$.  
There is (almost) no redshift dependence in the figure.

The dependence on the local environment seems to make sense. Overdense regions typically have young, spiral galaxies which have high $q$ and hence low RMS ellipticity. 
On the contrary, the underdense regions typically have old, elliptical galaxies and thus low $q$ and high RMS ellipticity.

From Figs.~\ref{fig:redshift_fluxlimited} and~\ref{fig:redshift_vollimited},
the region $0.4\le z < 0.55$ show signs of being marginally underdense but have low RMS ellipticities too that agree with the rest of the overdense regions.

When the B-band luminosity evolution is taken into account in selecting the sample, a systematic increase in the ellipticity at lower redshifts can be observed. We plot these alongside with stellar-mass selected samples, where a similar trend is found, in Fig.~\ref{fig:rms_ellip_Bband_mass}. 
Also, the RMS values of the ellipticites calculated from the second moments within each redshift bin are given in Fig.~\ref{fig:rms_ellip_momentbased}.

>>>>

To quantify galaxy morphology and shape, we will use several
quantities from the above fits.  First, from the single \sersic
profile fits, we use the \sersic index and the axis ratio.  The axis
ratio can also be used to derive a distortion,
\begin{equation}\label{eq:distortion}
e = \frac{1-q^2}{1+q^2}
\end{equation}
or an ellipticity,
\begin{equation}\label{eq:ellipticity}
\varepsilon = \frac{1-q}{1+q}.
\end{equation}

@@

We can test the influence of environment on the galaxy shapes by
comparing the distributions of the axis ratios for the overdense and
underdense redshift slices, or by encapsulating that distribution as a
single number, the RMS (root mean squared) ellipticity. By
volume-limiting the sample, we have avoided issues
wherein the flux limit leads to artificial changes in the sample as a
function of redshift.  We will also carry out tests to differentiate
between environmental effects versus evolution of the population with
redshift (at fixed mass).

%Then, we proceed to consider the root mean square ellipticity, a statistic that can characterize the shape of the population/sample of galaxies. 
\subsubsection{Comparing distributions}\label{subsubsec:compdist}
\begin{figure*}
 \centering
 \includegraphics[width=2.2\columnwidth]{axis_ratio_all}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_Bbandevolution.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_masscut.eps}
 \caption{The distributions of axis ratios of galaxies in \emph{all}
   overdense (OD) and \emph{all} underdense (UD) regions in the case
   of the luminosity-selected sample \s1 (left), luminosity-selected
   sample with $B$-band evolution taken into account \s2 (center), and
   thestellar-mass-selected sample \s3 (right). The upper panels show
   the histograms, 
 and the bottom panels show the cumulative distribution functions
 (CDF). The $p$-values computed using these CDFs are shown in Table. \ref{table:pvalues_all}.}
 \label{fig:axisratio_all}
\end{figure*}

@@

\subsubsection{RMS ellipticites}
\begin{figure}
 %\centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip1_noevolution.eps} \
% \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \caption{RMS distortions as a function of redshift. The horizontal errorbars indicate
          the width of the redshift bin, while the vertical ones
          are $1\sigma$ errorbars obtained by
          bootstrapping. The colors and shapes of the points indicate their
          environmental classification, as shown in the legend.  \rachel{The title and figure clearly show two
          samples.  The caption needs to explain which are solid
          points and which are open. Also, the horizontal axis label should
          be either $z$ or ``Photometric redshift ($z$)'', not ``Photo
          $z$'' (which is not a phrase we have used throughout the
          paper).  The change in horizontal axis label is needed for
          the other RMS ellipticity plots and anywhere else that it
          shows up in the paper.}}
 \label{fig:rms_ellip}
\end{figure}

\begin{figure}
 %\centering
  \includegraphics[width=1.0\columnwidth]{rms_ellip1_Bbandevolution.eps} \
%\includegraphics[width=1.0\columnwidth]{rms_ellip2_Bbandevolution.eps} \\
 \caption{RMS distortions as a function of redshift. The solid
   points correspond to the sample where the luminosity cut evolves by
   $-1.23$ magnitudes per unit redshift (\s$2$) and the open
          points correspond to the sample obtained from stellar mass
          cuts (\s$3$). \rachel{If I compare this with
            Fig.~\ref{fig:rms_ellip} I would have concluded that the
            samples were the opposite of what is stated here.  The solid points appear the
            same in both plots, so I thought that was the stellar
            mass-selected sample.  Please check and confirm which way
            is correct, and include the information in both captions.}}
 \label{fig:rms_ellip_Bband_mass}
\end{figure}

\begin{figure}
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_noevolution.eps} \
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_Bbandevolution_masscut.eps} \\
 \caption{RMS distortions as a function of redshift, with points
   defined in a similar way as in Fig.~\ref{fig:rms_ellip}.  In this
   case, the distortions are the moments-based shape estimates, rather
   than from fits to \sersic profiles. }
 \label{fig:rms_ellip_momentbased}         
\end{figure}

\rachel{Overall, I find it quite confusing that the underdense regions
  are red and the overdense regions are blue on all of these plots.
  The underdense regions are dominated by blue galaxies and the
  overdense are dominated by red galaxies, so having the plot colors
  reversed creates some cognitive dissonance.  Can you reverse them to
  more naturally match up point color with galaxy type?}

%The shape of a population of galaxies can be characterized by a single number called `RMS ellipticity'. If $q=b/a$ is the axis-ratio, then the ellipticity maybe defined as $\frac{1-q}{1+q}$
%or as $\frac{1-q^2}{1+q^2}$. The root mean squared (RMS) of ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}.
We can also carry out tests on a single statistic of the galaxy shape
distribution in each redshift slice, like the RMS ellipticity.  While
tests of a single quantity may seem less powerful than tests that use
the entire shape distributions, the advantage is that instead of
picking out pairs of redshift slices for our tests, we can easily compute our
statistic of interest for every single redshift slice, and look for
trends with both redshift and environment.

For the luminosity-selected sample without any evolution (\s1), the
RMS distortions (Eq.~\ref{eq:distortion}) of galaxies in each redshift bin are shown
in Fig.~\ref{fig:rms_ellip}.  (\rachel{The figure seems to show two
  samples, not just \s1.  Update the text to reflect this?}) For this and all similar figures, the
colors of the points were selected to easily differentiate between
galaxies in overdense, neutral, and underdense redshift slices.
As shown, the underdense regions
have  higher values for RMS ellipticities when compared to the
overdense regions. The difference between the underdense and overdense
regions for $z>0.5$ is significantly larger than any redshift
evolution across the $z>0.5$ range.  Our conclusions are very similar
if we use the RMS ellipticity from Eq.~\eqref{eq:ellipticity} instead
of the distortion. (\rachel{Our statements at the low redshift end
  regarding the amount of redshift evolution will depend on whether
  we're using the solid or open points, and the text and caption do
  not say which samples they correspond to, so I was not able to
  address this in the text.  Please do that once it is your turn to edit.})

The sign of the dependence on the local environment is reasonable when
compared with previous work on the morphology-density
relation. Overdense regions typically contain many old, elliptical
galaxies which are close to round (large axis ratio and low RMS ellipticity). 
In contrast, the underdense regions typically contain a larger
population of younger, disk galaxies, which have lower axis ratios and
higher RMS ellipticity. \rachel{Important note: this argument was
  given completely backwards.  I fixed it, but please watch out for
  this in future since it's a key point behind this work.}

From Figs.~\ref{fig:redshift_fluxlimited} and~\ref{fig:redshift_vollimited},
the $0.4\le z < 0.55$ redshift range shows signs of being marginally
underdense, but has somewhat low RMS ellipticities that agree with the rest of the overdense regions.

When the $B$-band luminosity evolution is taken into account in
selecting the sample, a systematic increase in the ellipticity at
lower redshifts can be observed. We plot these along with stellar-mass
selected samples, where a similar trend is found, in
Fig.~\ref{fig:rms_ellip_Bband_mass}.

Finally, we show an analogous plot of RMS ellipticities for all three
volume-limited sample using the moments-based shape estimates in
Fig.~\ref{fig:rms_ellip_momentbased}. The conclusions are quite
similar to those using the shapes from the \sersic profile fits, with
the underdense regions standing out in having larger RMS distortions 
than the other redshift slices, an effect that is substantially larger
than any average redshift evolution of the RMS distortions.

However, it is generally the case that the statistical significance of
trends in this section using a single statistic of the shape
distribution (the RMS) is less than the significance of the trends
seen using the entire axis ratio distributions in Sec.~\ref{subsubsec:compdist}.
